---
title: "R Structures"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


## Arithmetic and algebra

Let's start with the same arithmetic we were just doing in the Console. Notice that the math is surrounded by some special characters above and below, and has a grey background. This is RStudio's way of showing you the difference between an area to write text (here!) and an area where code is stored (called a code chunk).

```{r}
5 + 5
4 - 1
1 * 2
4 ^ 2
```

Try executing this chunk by clicking the green *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. These two actions do the same thing-- they run all the lines of code in a chunk, and show you all the output.

You can also run just one line at a time by putting your cursor on a line and pressing *Cmd+Enter*. Then, you'll just see the output for the one line. 

We can do the same with the algebra we just saw,

```{r}
a <- 1
b <- 2
a + b
A <- 3
a + b - A
```

What happened in your Environment when you ran that code? 

## Functions 

```{r}
round(3.1415)
factorial(3)
sqrt(9)

factorial(round(2.0015) + 1)
```


## Code Chunk Options 

There are a lot of things you can do in a code chunk: you can produce text output, tables, or graphics. 

You also have control over all these output via chunk options, which can be provided inside the curly braces (between ```{r and }). 

`echo=FALSE` indicates that the code will not be shown in the final document (though any results/output would still be displayed).

```{r, echo=FALSE}
x <- rnorm(100)
y <- 2*x + rnorm(100)
cor(x, y)
```
You use `results="hide"` to hide the results/output (but here the code would still be displayed).

```{r, results="hide"}
x <- rnorm(100)
y <- 2*x + rnorm(100)
cor(x, y)
```
You use `include=FALSE` to have the chunk *evaluated*, but neither the code nor its output displayed.

```{r, include=FALSE}
x <- rnorm(100)
y <- 2*x + rnorm(100)
cor(x, y)
```

For figures, you’ll want to use options like fig.width and fig.height

```{r scatterplot, fig.width=8, fig.height=6}
plot(x,y)
```

## R Markdown

Markdown provides an easy way to make standard types of formatted text, like

- *italics*
- **bold**
- `code`
- [links](rmarkdown.rstudio.com)
- Latex equations, $E = mc^{2}$
- etc.


## "knitted" HTML version

This functionality is supported by an R package called `knitr`, so we often talk about "knitted" versions of RMarkdown and R Notebook documents. 

When you save the notebook, an HTML file containing the code and output will be saved alongside it. This makes a nice, polished report of your work to share.

Click the *Preview* button at the top of this document or press *Cmd+Shift+K* (*Ctrl+Shift+K* on Windows) to preview the HTML file. Try clicking *Preview* now.

## R Objects and functions

```{r objects and functions}
the_answer <- 42
more_pi <- round(3.1415) + 1
factorial(more_pi)
```

## Vectors and matrices

```{r vectors}
vec <- c(1, 2, 3, 10, 100)
vec
```

## Data types

Numeric
```{r numeric}
1 + 1
3000000
class(0.00001)
```

Character
```{r character}
"hello"
class("hello")
```

```{r, error=TRUE}
"hello" + "world"
```

```{r}
nchar("hello")
?nchar

paste("hello", "world")
?paste
```

Logical
```{r logical}
3 > 4
class(TRUE)
class(T)
```

Factor (danger zone!)
```{r factor}
fac <- factor(c("a", "b", "c"))
fac
class(fac)
```

## Your turn 

Make a vector that contains the number 1, the letter R, and the logical TRUE

```{r}

```

## Manual Coercion
```{r}
#Convert character to a numeric
as.numeric("1")

#Convert a logical to character
as.character(TRUE)

```

## Data frames
```{r}
df <- data.frame(c(1, 2, 3), 
                 c("R","S","T"), 
                 c(TRUE, FALSE, TRUE))
class(df)
```
## str()

Run `str()` to see the structure of an object

```{r}
?str()
```

## Naming 
```{r}
nvec <- c(one = 1, two = 2, three = 3)
```

```{r}
ndf <- data.frame(numbers = c(1, 2, 3), 
                  letters = c("R","S","T"), 
                  logic = c(TRUE, FALSE, TRUE))
```

```{r}
names(ndf)
names(nvec)
names(df)
```
## Tibbles

```{r tibble}
df_tibble <- tibble(c(1, 2, 3), 
                    c("R","S","T"), 
                    c(TRUE, FALSE, TRUE))

#When running the same code, tibble will correctly store the second column as a character, not a factor. 
str(df_tibble)
```

## Subsetting

## The dplyr way
To use the dplyr package, we need to load it.

```{r}
library(dplyr)
```

## Beatles data
Let's make a toy data set to play with. 

```{r}
beatles <- data.frame(name = c("John", "Paul", "George", "Ringo"),
                                birth = c(1940, 1942, 1943, 1940), 
                                instrument = c("guitar", "bass", "guitar", "drums")
                                )
```

Create the same toy data set but in a tibble format. We can use the `tibble()` or the `tribble()` function

```{r}
beatles_tibble1 <- tibble(name = c("John", "Paul", "George", "Ringo"),
                          birth = c(1940, 1942, 1943, 1940), 
                          instrument = c("guitar", "bass", "guitar", "drums")
                        )


beatles_tibble2 <- tribble(
  ~name, ~birth, ~instrument, 
  #--|--|----
  "John", 1940, "guitar", 
  "Paul", 1942, "bass", 
  "George", 1943, "guitar", 
  "Ringo", 1940, "drums"
  )
```

## Select
Select is a way to extract columns from dataframes

```{r}
select(beatles, name, birth)
```

## Your turn
Select just the instrument column
```{r}
select(beatles, instrument)
```

## Filter

```{r}
filter(beatles, name == "George")
```

## Your turn
Alter the code to filter out just rows where birth is 1940:
```{r your turn filter}

```
## Logical comparisons

```{r Logical comparison basics}
1 < 3
1 > 3
c(1, 2, 3, 4, 5) > 3

```

```{r Logical comparison basics2 }
1 %in% c(1, 2, 3, 4)
1 %in% c(2, 3, 4)
c(3,4,5,6) %in% c(2, 3, 4)
```
## Your Turn Logical Comparions
```{r}
x <- c(1, 2, 3, 4, 5)


```


## More than one condition
```{r}
filter(beatles, birth==1940, instrument == "guitar")
filter(beatles, birth==1940 & instrument == "guitar")
```

## Your turn
Modify the code below to filter out the rows for which birth is 1943 or instrument is drums

```{r}
filter(beatles, birth==1940 & instrument == "guitar")
```
## Arrange
Sort rows based on specified column’s values. 

```{r}
arrange(beatles, birth)
```

You can sort by more than one column. Notice how the order changes when we add instrument as a column to sort by. 

```{r}
arrange(beatles, birth, instrument)
```
## Desc()
You can sort by descending order using `desc()`
```{r}
arrange(beatles, desc(birth))
```


## Baby names data
Names with at least 5 uses in a particular year, from the Social Security Administration. 

```{r Load babynames}
library(babynames)
```

## Your turn
Do you remember how to view a dataset?

```{r view babynames}
#Enter code to view a data set
```
## Your turn: return rows with your name 
```{r}
#Enter code to return only the babies with your first name. 
```

#Your turn: return all of the names where prop is greater than or equal to 0.08, All of the children named “Sea”, All of the names that have a missing value for n

```{r}

```
# Your turn: Use Boolean operators to alter the code below to return only the  rows that contain:
- Girls named Sea
- Names that were used by exactly 5 or 6 children in 1880
- Names that are one of Acura, Lexus, or Yugo

```{r}

```

## Steps

```{r}
boys_2015 <- filter(babynames, year == 2015, sex == "M")
boys_2015 <- select(boys_2015, name, n)
boys_2015 <- arrange(boys_2015, desc(n))
boys_2015
```

Nested nonsense

```{r}
arrange(select(filter(babynames, year == 2015, 
  sex == "M"), name, n), desc(n))
```

## Piping
The `%>%` operator is the pipe. We can reorganize our code like so:

```{r}
babynames %>%
  filter(year == 2015, sex == "M") %>%
  select(name, n) %>%
  arrange(desc(n))
```

[Tip: type Cmd Shift M to insert a pipe]

## Your turn
Use %>% to write a sequence of functions that: 
- Filter babynames to just the girls that were born in 2015
- Select the `name` and `n` columns
- Arrange the results so that the most popular names are near the top.

```{r}

```

## Summarise() 
Use summarise/summarize to compute a table of summaries

```{r}
babynames %>% summarise(total = sum(n), 
                        max = max(n))
```

## Your Turn

Use summarise() to compute three statistics about the data:

1. The first (minimum) year in the dataset  
2. The last (maximum) year in the dataset  
3. The total number of children represented in the data

```{r}


```

## Your Turn

Extract the rows where `name == "Khaleesi"`. Then use `summarise()` and a summary functions to find:

1. The total number of children named Khaleesi
2. The first year Khaleesi appeared in the data

## n() 
A function that calculates the number of rows in a dataset/group

```{r}
babynames %>% 
  summarize(n = n())

?n()
```

## n_distinct() 
The number of distinct values in a variable

```{r}
babynames %>% 
  summarise(n = n(), 
            distinct_names = n_distinct(name))
```
## Grouping cases

## Your Turn: Read in the pollution data set
```{r}
#Choose to read pollution data via a tribble

pollution <- tribble(
       ~city, ~size, ~amount,
  "New York", "large",      23,
  "New York", "small",      14,
    "London", "large",      22,
    "London", "small",      16,
   "Beijing", "large",     121,
   "Beijing", "small",     121
)

#The pollution set is also stored as a csv, we'll learn to use read_csv later. For now, just run the code and we'll explain what is happening later. 

pollution <- read_csv(here::here("practice_data", "pollution.csv"))
```
## Calculate the mean, sum, and n size of the pollutio data set
```{r}
pollution %>%
summarise(mean = mean(amount), 
          sum = sum(amount), 
          n = n())
```
## Group by
Groups cases by common values of one or more columns.

```{r}
pollution %>%  
  group_by(city)  %>%
  summarise(mean = mean(amount), 
            sum = sum(amount), 
            n = n())

```
Running `group_by()` by itself doesn't change the output of the data frame.  

```{r}
babynames %>% group_by(sex)
```
## Calculate the total names in the `babynames` data set by sex:  

```{r}
babynames %>%  
  group_by(sex) %>%
  summarise(total = sum(n))

```
## Your Turn

Use `group_by()`, `summarise()`, and `arrange()` to display the ten most popular names. Compute popularity as the total number of children of a single gender given a name.

```{r}


```
## Your turn

Use grouping to calculate and then plot the number of children born each year over time.
```{r}
babynames %>%
    group_by(year) %>%  
    summarise(n_children = sum(n)) %>%  
    ggplot() +
    geom_line(mapping = aes(x = year, y = n_children))

```

## Mutate

Create new columns.

```{r}
babynames %>%
  mutate(percent = round(prop*100, 2),
         nper = round(percent))
```
## Your turn

Use min_rank() and mutate() to rank each row in  babynames from largest n to lowest n.
```{r}

```

## Your turn

- Compute each name's rank within its year and sex.
- Then compute the median rank for each combination of name and sex, and arrange the results from highest  median rank to lowest.

```{r}

```

## Joining Data sets 

## Your turn: Read in toy data sets band and instrument
```{r}
#Create the data set using tribble
band <- tribble(
    ~name, ~band, 
    "Mick", "Stones", 
    "John", "Beatles", 
    "Paul", "Beatles"
)

instrument <- tribble(
  ~name, ~plays, 
  "John", "guitar",
  "Paul", "bass",
  "Keith", "guitar"
)

# Read in using readr
band <- read_csv(here::here("practice_data", "band.csv"))
instrument <- read_csv(here::here("practice_data", "instrument.csv"))
```
## Types of joins

```{r}
band %>% left_join(instrument, by = "name")
band %>% right_join(instrument, by = "name")
band %>% full_join(instrument, by = "name")
band %>% inner_join(instrument, by = "name")
```


## NYC Flights Package
Data about every flight that departed La  Guardia, JFK, or Newark airports in 2013

```{r}
install.packages("nycflights13")
library(nycflights13)
```


```{r}
View(flights)
```

```{r}
View(flights %>% select("carrier"))
```

```{r}
View(airlines)
```

## Your turn
Which airlines had the largest arrival delays? Work in groups to complete the code below. 
```{r}
flights %>% 
  drop_na(arr_delay) %>%
  #_________ %>%
  group_by(name) %>%
  #__________ %>%
  #arrange(______)
```

## Your turn
Read in the toy dataset instrument2

```{r}
instrument2 <- read_csv(here::here("practice_data", "instrument2.csv"))
```

## What if the names don't match?

```{r}
band %>% left_join(instrument2, by = c("name" = "artist"))
```


```{r}
airports %>% select(1:3)
```
```{r}
flights %>% select(14:15)
```


```{r}
airports %>% left_join(flights, by = c("faa" = "dest"))
```

## Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  

* Make tables of summaries with `summarise()`  
* Make new variables, with `mutate()`  
* Do groupwise operations with `group_by()`

* Connect operations with `%>%`  

* Use `left_join()`, `right_join()`, `full_join()`, or `inner_join()` to join datasets
* Use `semi_join()` or `anti_join()` to filter datasets against each other

## Gather

## Read in the toy data
```{r}
# Toy data
cases <- tribble(
  ~Country, ~"2011", ~"2012", ~"2013",
      "FR",    7000,    6900,    7000,
      "DE",    5800,    6000,    6200,
      "US",   15000,   14000,   13000
)

pollution <- tribble(
       ~city, ~size, ~amount,
  "New York", "large",      23,
  "New York", "small",      14,
    "London", "large",      22,
    "London", "small",      16,
   "Beijing", "large",     121,
   "Beijing", "small",     121
)

x <- tribble(
  ~x1, ~x2,
  "A",   1,
  "B",  NA,
  "C",  NA,
  "D",   3,
  "E",  NA
)

# To avoid a distracting detail during class
names(who) <- stringr::str_replace(names(who), "newrel", "new_rel")
```

```{r}
cases
```
## Your turn

On a sheet of paper, draw how the cases data set would look if it had the same values grouped into three columns: **country**, **year**, **n**


## Gather

```{r}
cases %>% gather(key = "year", value = "n", 2:4)
cases %>% gather("year","n", 2:4)
cases %>% gather("year","n", "2011", "2012", "2013")
cases %>% gather("year","n", -Country)
```

## Your turn

Use `gather()` to reorganize `table4a` into three columns: **country**, **year**, and **cases**.

```{r}

```
## Your turn

Use `gather()` to reorganize `table4a` into three columns: **country**, **year**, and **cases**.

```{r}

```

## Your turn

On a sheet of paper, draw how this data set would look if it had the same values grouped into three columns: **city**, **large**, **small**


